---
title: "Basic Analysis chevron"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE}
library(ggplot2)
library(tidyr)
library(gridExtra)
```

# News sentiment score for chevron (factor we want to investigate)
```{r echo=FALSE}
news_sentiment <- read.csv("Basic_analysis_data/ave-daily-news-sentiment-chevron-2020.csv")
colnames(news_sentiment)[1] <- "date"

# remove the time from news date
news_sentiment$date <- gsub(" 0:00", "", news_sentiment$date)
news_sentiment$date <- gsub("/", "-", news_sentiment$date)

#head(news_sentiment)
```

```{r echo=FALSE}
# uniform time format
news_sentiment$date <- format(as.Date(news_sentiment$date, "%m-%d-%Y"), '%Y-%m-%d')
news_sentiment$date <- as.Date(news_sentiment$date)
```

```{r echo=FALSE}
par(mfrow=c(1,2))
hist(news_sentiment$Daily.News.Sentiment, xlab = 'sentiment score', main = '')
plot(news_sentiment$Daily.News.Sentiment ~ as.Date(news_sentiment$date), xlab = 'date',
     ylab = 'sentiment score')
abline(h=-0.2, col = 'red')
```

# NYSE stock closing price for chevron (target variable)
```{r echo=FALSE, message=FALSE}
NYSE_stock <- 
  read.csv("Basic_analysis_data/nyse-energy-company-stock-closing-prices-2020.csv")
colnames(NYSE_stock)[1] <- "Stock Price"
NYSE_stock$Date <- as.Date(NYSE_stock$Date)
NYSE_stock <- subset(NYSE_stock, select = -c(Super.Region, Ticker, Time.Level, Source, Sub.Sector, Stock.Attribute, Currency, Data.Provider, Stock.Exchange, Location))
chevron_stock <- NYSE_stock[which(NYSE_stock$Stock.Name == "CHEVRON CORPORATION"),]
```

## Basic summary of chevron stock price
```{r echo=FALSE}
par(mfrow=c(1,2))
hist(chevron_stock$`Stock Price`, main = '', xlab = 'chevron Stock Price')
plot(chevron_stock$`Stock Price` ~ as.Date(chevron_stock$Date), xlab = 'Date', ylab = 'chevron stock price')
```

# Crude oil commodity
```{r echo=FALSE}
Crude_oil_c <- read.csv("Basic_analysis_data/crude-oil-commodity-prices-2020.csv")
Crude_oil_c <- Crude_oil_c[Crude_oil_c$Indicator=="SETTLE",]
colnames(Crude_oil_c)[1] <- "Price"
Crude_oil_c$Date <- as.Date(Crude_oil_c$Date)
```

## Basic summary for Ice Brent (settle)
```{r echo=FALSE}
ice_brent <- Crude_oil_c[Crude_oil_c$Commodity.And.Exchange == "ICE BRENT CRUDE OIL FUTURES",]
```

```{r echo=FALSE}
par(mfrow=c(1,2))
hist(ice_brent$Price, xlab = 'price', main = "Ice Brent")
plot(ice_brent$Price ~ as.Date(ice_brent$Date), xlab = 'date', ylab = 'price',
     main = 'Ice Brent')
```

## Basic summary for NYMEX 
```{r echo=FALSE}
NYMEX <- Crude_oil_c[Crude_oil_c$Commodity.And.Exchange == "NYMEX CRUDE OIL FUTURES",]
```

```{r echo=FALSE}
par(mfrow=c(1,2))
hist(NYMEX$Price, main = 'NYMEX', xlab = 'Price')
plot(NYMEX$Price ~ as.Date(NYMEX$Date), xlab = 'date', ylab = 'price', main = 'NYMEX')
```

## Compare Ice Brent and NYMEX
```{r echo=FALSE}
match.date <- intersect(NYMEX$Date, ice_brent$Date)
plot(NYMEX[NYMEX$Date %in% match.date,]$Price, ice_brent[ice_brent$Date %in% match.date,]$Price,
     xlab = 'NYMEX', ylab = 'Ice Brent', main = 'Ice Brent vs. NYMEX')
```

One should clearly observe that the trend of crude oil commodity and chevron stock price are consistent.  
```{r echo=FALSE}
par(mfrow=c(1,3))
plot(chevron_stock$`Stock Price` ~ as.Date(chevron_stock$Date), xlab = 'Date', ylab = 'chevron stock price',
     main = 'chevron')
plot(NYMEX[NYMEX$Price>0,]$Price ~ as.Date(NYMEX[NYMEX$Price>0,]$Date), 
     xlab = 'date', ylab = 'price', main = 'NYMEX (negative price removed)')
plot(ice_brent$Price ~ as.Date(ice_brent$Date), xlab = 'date', ylab = 'price',
     main = 'Ice Brent')
```

# Stock Market Indicator
```{r echo=FALSE}
mkt_indicator <- read.csv("Basic_analysis_data/some-stock-market-indicators-&-index-2020.csv")
mkt_indicator$Date <- as.Date(mkt_indicator$Date)
colnames(mkt_indicator)[1] <- "Price"
SP500 <- mkt_indicator[mkt_indicator$Description == 'S&P 500',]
Dow_Jones <- mkt_indicator[mkt_indicator$Description != 'S&P 500',]
```

## Summary of Dow Jones

```{r echo=FALSE, message=FALSE}
ggplot(Dow_Jones, aes(x = Price, fill = Description)) + 
  geom_histogram(position = "identity", alpha = 0.4)
```

It appears different Dow Jones indexes are quite different.  

```{r echo=FALSE, message=FALSE}
composite_ave <- Dow_Jones[Dow_Jones$Description=='DOW JONES COMPOSITE AVERAGE',]
indust_ave <- Dow_Jones[Dow_Jones$Description=='DOW JONES INDUSTRIAL AVERAGE',]
transp_ave <- Dow_Jones[Dow_Jones$Description=='DOW JONES TRANSPORTATION AVERAGE',]
utility_ave <- Dow_Jones[Dow_Jones$Description=='DOW JONES UTILITY AVERAGE',] 
```

```{r echo=FALSE, message=FALSE}
p1 <- ggplot(composite_ave, aes(x = Price))+ geom_histogram() + labs(title = 'COMPOSITE AVERAGE')
p2 <- ggplot(indust_ave, aes(x = Price))+ geom_histogram() + labs(title = 'INDUSTRIAL AVERAGE')
p3 <- ggplot(transp_ave, aes(x = Price))+ geom_histogram() + labs(title = 'TRANSPORTATION AVERAGE')
p4 <- ggplot(utility_ave, aes(x = Price))+ geom_histogram() + labs(title = 'UTILITY AVERAGE')
grid.arrange(p1, p2, p3, p4, nrow = 2, ncol = 2)
```


```{r echo=FALSE}
p1 <- ggplot(composite_ave, aes(x=Date, y=Price)) + geom_line() + labs(title = 'COMPOSITE AVERAGE')
p2 <- ggplot(indust_ave, aes(x=Date, y=Price)) + geom_line() + labs(title = 'INDUSTRIAL AVERAGE')
p3 <- ggplot(transp_ave, aes(x=Date, y=Price)) + geom_line() + labs(title = 'TRANSPORTATION AVERAGE')
p4 <- ggplot(utility_ave, aes(x=Date, y=Price)) + geom_line() + labs(title = 'UTILITY AVERAGE')
grid.arrange(p1, p2, p3, p4, nrow = 2, ncol = 2)
```

## Summary of S&P 500
```{r echo=FALSE, message=FALSE}
sp1 <- ggplot(data = SP500, aes(x = Price)) + geom_histogram() + labs(title = 'S&P 500')
sp2 <- ggplot(SP500, aes(x = Date, y = Price)) + geom_line() + labs(title = 'S&P 500')
grid.arrange(sp1, sp2, nrow = 1, ncol = 2)
```


# Apple Mobility California
```{r echo=FALSE}
library(readxl)
apple <- read_excel("Basic_analysis_data/apple-mobility-california-2020.xlsx")
apple$Date <- as.Date(apple$Date)
```

```{r echo=FALSE}
#unique(apple$`Transportation Type`)
apple_transit <- apple[apple$`Transportation Type`=='TRANSIT',]
apple_driving <- apple[apple$`Transportation Type`=='DRIVING',]
apple_walking <- apple[apple$`Transportation Type`=='WALKING',]
```

## Basic Summary
```{r echo=FALSE, message=FALSE}
ap1 <- ggplot(apple, aes(x = Value, fill = `Transportation Type`)) + 
  geom_histogram(position = "identity", alpha = 0.5) + labs(title = 'Apple Mobility')
ap1

ap2 <- ggplot(apple, aes(x = Date, y = Value, color = `Transportation Type`)) + 
  geom_point(position = "identity", alpha = 0.7) + labs(title = 'Apple Mobility')
ap2
#ap2 <- ggplot(apple_driving, aes(x = Value)) + geom_histogram() + labs(title = 'Apple Driving Mobility')
#ap3 <- ggplot(apple_transit, aes(x = Value)) + geom_histogram() + labs(title = 'Apple Transit Mobility')
#ap4 <- ggplot(apple_walking, aes(x = Value)) + geom_histogram() + labs(title = 'Apple Waling Mobility')
#grid.arrange(ap1, ap2, ap3, ap4, nrow = 2, ncol = 2)
```

# Google Mobility California LA
```{r echo=FALSE}
#google_mob <- read.csv("C:/Users/yuche/Desktop/UCF-Math/2021 DS Comp/google-mobility-index-la-county-2020.csv")
#google_mob$Date <- gsub("/", "-", google_mob$Date)
#google_mob$Date <- format(as.Date(google_mob$Date, "%m-%d-%Y"), '%Y-%m-%d')
#colnames(google_mob)[1] <- "Value"
```

## Summary
```{r echo=FALSE, message=FALSE}
#gp1 <- ggplot(google_mob, aes(x = Value, fill = Indicator)) + 
 # geom_histogram(position = "identity", alpha = 0.5) + labs(title = 'Google Mobility')
#gp1+ theme(legend.position="bottom", legend.text = element_text(size = 6)) + guides(fill = guide_legend(nrow = 3))

#gp2 <- ggplot(google_mob, aes(x = Date, y = Value)) + 
 # geom_point(aes(color = Indicator)) + labs(title = 'Google Mobility')
#gp2 + theme(legend.position="bottom", legend.text = element_text(size = 6)) + guides(colour = guide_legend(nrow = 3))
```

# Investigate Models
Note that we used previous day's news sentiment vs against the current day's other features; the reason is because the current day's mews sentiment shows no correlation at all with respect to current day's stock price. This could be because some news are coming out after the trading center closed; or it could be because the effect of news on the market is delayed.  

```{r echo=FALSE}
common_date <- Reduce(intersect, list(news_sentiment$date, chevron_stock$Date, ice_brent$Date,
                                      NYMEX$Date, composite_ave$Date, indust_ave$Date,
                                      transp_ave$Date, utility_ave$Date, SP500$Date,
                                      apple_driving$Date, apple_transit$Date, apple_walking$Date))
```

```{r echo=FALSE}
date <- news_sentiment[news_sentiment$date %in% common_date,]$date
chevron_stock_price <- chevron_stock[chevron_stock$Date %in% common_date,]$`Stock Price`
news_sentiment_score <- news_sentiment[news_sentiment$date %in% common_date,]$Daily.News.Sentiment
ice_brent_settle <- ice_brent[ice_brent$Date %in% common_date,]$Price
NYMEX_settle <- NYMEX[NYMEX$Date %in% common_date,]$Price
DJ_comp <- composite_ave[composite_ave$Date %in% common_date,]$Price
DJ_indus <- indust_ave[indust_ave$Date %in% common_date,]$Price
DJ_tran <- transp_ave[transp_ave$Date %in% common_date,]$Price
DJ_utility <- utility_ave[utility_ave$Date %in% common_date,]$Price
SP500_price <- SP500[SP500$Date %in% common_date,]$Price
apple_driving_index <- apple_driving[apple_driving$Date %in% common_date,]$Value
apple_transit_index <- apple_transit[apple_transit$Date %in% common_date,]$Value
apple_walking_index <- apple_walking[apple_walking$Date %in% common_date,]$Value

df <- data.frame(date, chevron_stock_price, news_sentiment_score, ice_brent_settle, NYMEX_settle, DJ_comp,
                 DJ_indus, DJ_tran, DJ_utility, SP500_price, apple_driving_index, apple_transit_index,
                 apple_walking_index)
```


```{r echo=FALSE}
#df.adj <- data.frame(df[-81,c(1:2, 4:13)], df[-1, 3])
#colnames(df.adj)[13] <- "news_sentiment_score"

#df.norm <- data.frame(df[,1], scale(df[,2]), df[,3], scale(df[,4:13]))
df.norm <- data.frame(df[,1], scale(df[,-1]))
colnames(df.norm) <- names(df)
```

```{r echo=FALSE, message=FALSE}
#attach(df.adj)
```

```{r echo=FALSE, message=FALSE}
cormat <- round(cor(df.norm[,-1]),2)
library(reshape2)
melted_cormat <- melt(cormat)
# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
  cormat[upper.tri(cormat)] <- NA
  return(cormat)
  }
# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)
  }
upper_tri <- get_upper_tri(cormat)

melted_cormat <- melt(upper_tri, na.rm = TRUE)

reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}

# Reorder the correlation matrix
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()

ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
```

We noticed that the news sentiment is somewhat correlated with chevron Stock price, Dow Jones Utility average and apple mobility transit index.  

## Investigate news sentiment:
```{r echo=FALSE}
lm.1 <- lm(chevron_stock_price ~ . - news_sentiment_score, data = df.norm[,-1])
lm.2 <- lm(chevron_stock_price ~ . , data = df.norm[,-1])
anova(lm.1, lm.2)
```

## Stepwise method selecting model:(AIC)
```{r echo=FALSE, message=FALSE}
library(MASS)

step.model <- stepAIC(lm.2, direction = "both", 
                      trace = FALSE)
summary(step.model)
```

```{r echo=FALSE}
step.news <- step(lm.2, scope = list(lower = ~ news_sentiment_score), direction = "both",
                  trace = FALSE)
summary(step.news)
```


## Stepwise CV:
```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(caret)
library(leaps)
```



```{r echo=FALSE, message=FALSE}
# Set seed for reproducibility
set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model2 <- train(chevron_stock_price ~ ., data = df.norm[,-1],
                    method = "leapBackward", 
                    tuneGrid = data.frame(nvmax = 1:11),
                    trControl = train.control
                    )


#coef(step.model$finalModel, as.numeric(step.model$bestTune))
coef(step.model2$finalModel, as.numeric(step.model2$bestTune))

#step.model$results
```

```{r}
lm.3 <- lm(chevron_stock_price ~ ice_brent_settle + DJ_tran + DJ_utility + SP500_price, 
           data = df.norm)
lm.4 <- lm(chevron_stock_price ~ ice_brent_settle + DJ_tran + DJ_utility + SP500_price + news_sentiment_score, 
           data = df.norm)
anova(lm.3, lm.4)
```


## Take a look at performance of Linear model selected by stepwise method
```{r echo=FALSE}
estimated_stock <- step.model$fitted.values
df.estimate <- data.frame(df, estimated_stock)
ggstep1 <- ggplot(df.estimate) + 
  geom_line(aes(x=date, y=estimated_stock, color = 'red')) + 
  geom_line(aes(x = date, y = scale(chevron_stock_price), color = 'blue')) + ylab("Price") + xlab('Date') +
  scale_color_discrete(name = "", labels = c("Real", "Estimate")) + labs(title = 'step wise AIC')

estimated_stock <- lm.3$fitted.values
df.estimate <- data.frame(df, estimated_stock)
ggstep2 <- ggplot(df.estimate) + 
  geom_line(aes(x=date, y=estimated_stock, color = 'red')) + 
  geom_line(aes(x = date, y = scale(chevron_stock_price), color = 'blue')) + ylab("Price") + xlab('Date') +
  scale_color_discrete(name = "", labels = c("Real", "Estimate")) + labs(title = 'stepwise CV')

grid.arrange(ggstep1, ggstep2, nrow = 2, ncol = 1)
```

```{r echo=FALSE}
#write.csv(df,"organized_chevron.csv", row.names = FALSE)
```


